---
title: "Day 21 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
options(scipen = 999)
```

## Part 1

```{r}
input<-read_lines("Day21Sample.txt")
```

Make functions that just do these things, bring in a vector, spit out a vector:

```{r}
### switches two letters based on their position
swapposition<-function(vec,x,y){
### because 1 based, need to add 1 to x & y
  x<-x+1
  y<-y+1
  z<-vec[x]
  vec[x]<-vec[y]
  vec[y]<-z
  vec}

### switches two letters based on their letter
swapletter<-function(vec,x,y){
  px<-which(vec==x)
  py<-which(vec==y)
  vec[px]<-y
  vec[py]<-x
  vec}

### moves to the left
rotateleft<-function(vec,x){
  if(x!=0){
    for(i in 1:x){
      vec<-c(vec[-1],vec[1])}}
  vec}

### move to the right
rotateright<-function(vec,x){
  if(x!=0){
    for(i in 1:x){
      vec<-c(tail(vec,1),vec[-length(vec)])}}
  vec}

### rotate based on position of 
rotatebased<-function(vec,x){
  px<-which(vec==x)
  ### because the instructions say to add 1, don't need to do that.
  if(px>=5){px<-px+1}
  ### rotate it that many times
  for(i in 1:px){
    vec<-c(tail(vec,1),vec[-length(vec)])}
  vec}

###reverse between
reversexy<-function(vec,x,y){
  x<-x+1
  y<-y+1
  vec[x:y]<-rev(vec[x:y])
  vec}

###move
movexy<-function(vec,x,y){
  x<-x+1
  y<-y+1
  z<-vec[x]
  vec<-vec[-x]
  if(y==1){vec<-c(z,vec)
  }else if(y>length(vec)){vec<-c(vec,z)
  }else(vec<-c(vec[1:y-1],z,vec[y:length(vec)]))
  vec}

```

Then the function that runs them all

```{r}
scramble<-function(instructions=input,v="abcde"){
  v<-unlist(str_split(v,""))
  for(i in 1:length(instructions)){
    inst<-unlist(str_split(instructions[i]," "))
    cat(inst,"\t")
    if(inst[1]=="swap"&&inst[2]=="position"){
      v<-swapposition(v,as.numeric(inst[3]),as.numeric(inst[6]))
    }else if(inst[1]=="swap"&&inst[2]=="letter"){
      v<-swapletter(v,inst[3],inst[6])
    }else if(inst[1]=="rotate"&&inst[2]=="right"){
      v<-rotateright(v,as.numeric(inst[3]))
    }else if(inst[1]=="rotate"&&inst[2]=="left"){
      v<-rotateleft(v,as.numeric(inst[3]))
    }else if(inst[1]=="rotate"&&inst[2]=="based"){
      v<-rotatebased(v,inst[7])
    }else if(inst[1]=="reverse"){
      v<-reversexy(v,as.numeric(inst[3]),as.numeric(inst[5]))
    }else if(inst[1]=="move"){
      v<-movexy(v,as.numeric(inst[3]),as.numeric(inst[6]))
    }else{cat("something went wrong\n")}
  cat(v,"\n")}
str_flatten(v)}
```


```{r}
scramble()
```


##Part 2
For the most part, it's just going to be switching minor things - swaps won't change, rotate the other way, move will go in reverse.  Rotate based on position will be slightly different - but it's mapped 1 to 1.  So, a reverse rotate function is needed.  This is very specific for 8 characters

```{r}
revrotate<-function(vec,x){
    px<-which(vec==x)
    switch(px,
           "1"=vec<-c(tail(vec,7),head(vec,1)),
           "2"=vec<-c(tail(vec,7),head(vec,1)),
           "3"=vec<-c(tail(vec,2),head(vec,6)),
           "4"=vec<-c(tail(vec,6),head(vec,2)),
           "5"=vec<-c(tail(vec,1),head(vec,7)),
           "6"=vec<-c(tail(vec,5),head(vec,3)),
           "7"=vec<-vec,
           "8"=vec<-c(tail(vec,4),head(vec,4)),
           cat("something reversed wrong\n"))
    vec}
  
```

Then unscramble:


```{r}
unscramble<-function(instructions=input,v="abcde"){
  v<-unlist(str_split(v,""))
  for(i in length(instructions):1){
    inst<-unlist(str_split(instructions[i]," "))
    cat(inst,"\t")
    ### swap position stays the same
    if(inst[1]=="swap"&&inst[2]=="position"){
      v<-swapposition(v,as.numeric(inst[3]),as.numeric(inst[6]))
      ### swap letter stays the same
    }else if(inst[1]=="swap"&&inst[2]=="letter"){
      v<-swapletter(v,inst[3],inst[6])
      ### if it says rotate right, rotate left
    }else if(inst[1]=="rotate"&&inst[2]=="right"){
      v<-rotateleft(v,as.numeric(inst[3]))
      ### if it says rotate left, rotate right
    }else if(inst[1]=="rotate"&&inst[2]=="left"){
      v<-rotateright(v,as.numeric(inst[3]))
      ### if it says "rotate based on" use the reverse rotate
    }else if(inst[1]=="rotate"&&inst[2]=="based"){
      v<-revrotate(v,inst[7])
      ### reverse stays the same
    }else if(inst[1]=="reverse"){
      v<-reversexy(v,as.numeric(inst[3]),as.numeric(inst[5]))
      ### move switches positions
    }else if(inst[1]=="move"){
      v<-movexy(v,as.numeric(inst[6]),as.numeric(inst[3]))
    }else{cat("something went wrong\n")}
  cat(v,"\n")}
str_flatten(v)}
```

Because the reverse only works for 8 characters, not including with my input
```{r,eval=FALSE}
unscramble(input,"abcde")
```

