---
title: "Day 25 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-read_lines("../../AoCData/AOC2016/Day25.txt")
```

Start with the assembunny code from day 23


```{r}
togit<-function(ti){
  ti<-unlist(str_split(ti," "))
  if(length(ti)==2){
   if(ti[1]=="inc"){ti[1]<-"dec"}else{ti[1]<-"inc"}
  }else{
    if(ti[1]=="jnz"){ti[1]<-"cpy"}else{ti[1]<-"jnz"}}
  ti<-str_flatten(ti,collapse=" ")
ti}

```

Make two minor changes - one to get an "out" and then to say how long to keep going

```{r}
assembunny<-function(instructions=input,a,b,c,d,outlength){
  ### set up registers
  r<-dict(list("a"=a,"b"=b,"c"=c,"d"=d))
  i<-1
  outline<-c()
  ### i is the instruction being read stop when out of instructions
  while(i<=length(instructions)&&length(outline)<outlength){
#    cat(instructions[i],"\n")
    ### break this into pieces in a list
    inst<-as.list(unlist(str_split(instructions[i]," ")))
    ### turn numbers into numbers rather than strings
    for(j in 2:length(inst)){
      if(str_detect(inst[[j]],"[0-9]")){
        inst[[j]]<-as.numeric(inst[[j]])
      }}
    ### run through the instructions
    switch(inst[[1]],
           ### if cpy, put the number in the register
           "cpy"={if(is.numeric(inst[[2]])){r$set(inst[[3]],inst[[2]])}else{r$set(inst[[3]],r$get(inst[[2]]))}
             i<-i+1},
           ### increase the register
           "inc"={r$set(inst[[2]],r$get(inst[[2]])+1)
             i<-i+1},
           ### decrease the register
           "dec"={r$set(inst[[2]],r$get(inst[[2]])-1)
             i<-i+1},
           ### jump if NOT zero (in my input, there is a placeholder instruction,
           ### so while the second slot is usually a register, it doesn't have to be)
           "jnz"={
             if(is.numeric(inst[[3]])){n3<-inst[[3]]}else{n3<-r$get(inst[[3]])}
             if(is.numeric(inst[[2]])){
             if(inst[[2]]!=0){i<-i+n3}else{i<-i+1}
             }else{if(r$get(inst[[2]])!=0){i<-i+n3}else{i<-i+1}}},
           ###  
           "tgl"={
             x<-inst[[2]]
             if(!is.numeric(x)){x<-r$get(x)}
             if((i+x)<=length(instructions)){
               instructions[i+x]<-togit(instructions[i+x])}
           i<-i+1
           },
           "out"={x<-inst[[2]]
           if(!is.numeric(x)){x<-r$get(x)}
           outline<-c(outline,x)
           i<-i+1},
           cat("something went wrong\n"))}
  list(r,outline)}
```

From there, guess and check?
I'm sure there's a better way - but this worked.

```{r}
sapply(1:20,function(x){y<-assembunny(input,x,0,0,0,1)[[2]]
y})
```
So, evens are possible

```{r}
sapply(seq(from=2,to=16,by=2),function(x){y<-assembunny(input,x,0,0,0,2)[[2]]
y})
```
Every 4, starting at 2 (from here, start checkng the next 4 - because it seems obvious that's how the repeat will go)
```{r}
sapply(seq(from=2,to=14,by=4),function(x){y<-assembunny(input,x,0,0,0,3)[[2]]
y})
```
Every 8, starting at 6 (from here, start checkng the next 4 - because it seems obvious that's how the repeat will go)
```{r}
sapply(seq(from=6,to=30,by=8),function(x){y<-assembunny(input,x,0,0,0,4)[[2]]
y})
```
Every 16, starting at 6
```{r}
sapply(seq(from=6,to=54,by=16),function(x){y<-assembunny(input,x,0,0,0,5)[[2]]
y})
```
Every 32, starting at 22
```{r}
sapply(seq(from=22,to=118,by=32),function(x){y<-assembunny(input,x,0,0,0,6)[[2]]
y})
```
Every 64, starting at 54
```{r}
sapply(seq(from=54,to=246,by=64),function(x){y<-assembunny(input,x,0,0,0,7)[[2]]
y})
```
Every 128, starting at 54
```{r}
sapply(seq(from=54,to=438,by=128),function(x){y<-assembunny(input,x,0,0,0,8)[[2]]
y})
```
Every 256, starting at 182
```{r}
sapply(seq(from=182,to=950,by=256),function(x){y<-assembunny(input,x,0,0,0,9)[[2]]
y})
```
Every 512, starting at 182
```{r}
sapply(seq(from=182,to=1718,by=512),function(x){y<-assembunny(input,x,0,0,0,10)[[2]]
y})
```

Every 1024, starting at 182
```{r}
sapply(seq(from=182,to=3524,by=1024),function(x){y<-assembunny(input,x,0,0,0,11)[[2]]
y})
```
Every 2048, starting at 182
```{r}
sapply(seq(from=182,to=6326,by=2048),function(x){y<-assembunny(input,x,0,0,0,12)[[2]]
y})
```

Every 4096, starting at 182
```{r}
sapply(seq(from=182,to=12470,by=4096),function(x){y<-assembunny(input,x,0,0,0,13)[[2]]
y})
```
Every 8192, starting at 182
```{r}
sapply(seq(from=182,to=24758,by=8192),function(x){y<-assembunny(input,x,0,0,0,15)[[2]]
y})
```
The pattern changes and all of a sudden, the first two examples work, the second two don't... So, run 182 for 32 and see if it "works"

```{r}
daytwentyfive<-assembunny(input,182,0,0,0,32)
daytwentyfive[[2]]
```
Which it does - so there's an answer worth guessing.