---
title: "Day 15 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
options(scipen = 999)
```

```{r}
input<-read_lines("Day15Sample.txt")
### pull out info, so that it's disc number, positions, starting position

discs<-lapply(input,function(x){
  x<-as.numeric(unlist(str_split(x,"[# .]"))[c(3,5,13)])})
```

## Part 1

Quick function that returns true if it drops through, false if it would bounce

```{r}
passthrough<-function(dt,dsc){
  if((dt+dsc[1]+dsc[3])%%dsc[2]==0){return(TRUE)}else{return(FALSE)}}
```

Find the first place for the ball to fall through the first disc:

```{r}
part1<-0
bounce<-TRUE
while(bounce){
  ok<-passthrough(part1,discs[[1]])
  if(ok){bounce<-FALSE}else{part1<-part1+1}}
part1
```

For all the other discs (because the number of slots is always prime):
start with the current shortest time - and the period where all the above discs will line up
check the next X cycles (where X is the number of slots in the current disc)
```{r}
d<-2
while(d<=length(discs)){
### poss(ible) stars with the current valid time - then repeats for each period 
    poss<-seq(from=part1,
            to=discs[[d]][2]*prod(sapply(discs[1:(d-1)],function(x){x[2]})),
            by=prod(sapply(discs[1:(d-1)],function(x){x[2]})))
    ###check each repeat to see if it will fall through this new disc - the one that does is the next starting point
    part1<-poss[which(sapply(poss,function(x){passthrough(x,discs[[d]])}))]
    d<-d+1}
part1
```

## Part 2

The new disc has 11 slots, so do the above with 11.

```{r}
####These are the possible answers
poss<-seq(from=part1,to=11*prod(sapply(discs,function(x){x[2]})),by=prod(sapply(discs,function(x){x[2]})))

### figure out which one will pass through the new disc (c(7,11,0))
part2<-poss[which(sapply(poss,function(x){passthrough(x,c(7,11,0))}))]
part2
```

To prove it works - quickly run it through all discs

```{r}
sapply(discs,function(x){passthrough(part2,x)})
```
