---
title: "Day 10 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-read_lines("Day10Sample.txt")
```

## Part 1
As with all of these, I'm going to read in line by line over and over and over again until I get an answer - for this one, stop as soon as I have a robot comparing the two values in my data:


This adds values for value adding steps and compares chips for the remainder, but stops when it gets the comparison that I'm looking for.
```{r}

microrunner<-function(instlist){
  d<-dict()
  ### as long as there are still instructions
  while(length(instlist)>0){
    ### start at the top of the instruction list
        i<-1
    while(i<=length(instlist)){
      ### get an instruction and split it in pieces
      inst<-instlist[i]
      inst<-unlist(str_split(inst," "))
      ### if it is in the form "value X goes to bot Y"
      if(length(inst)==6){
        ### add the value to the bot (sort so that the low value is first)
        adval<-str_flatten(c(inst[5],inst[6])," ")
        if(d$has(adval)){d$set(adval,sort(c(d$get(adval),as.numeric(inst[2]))))}else{d$set(adval,c(as.numeric(inst[2])))}
        ### remove that instruction from the instruction list
        instlist<-instlist[-i]
        next
      }else{
        ### the instruction is of the form: bot X gives low to Y and high to Z
        ### find bot X
        fr<-str_flatten(c(inst[1],inst[2])," ")
        ### if bot X doesn't have two chips, move on to the next instruction
        if(!d$has(fr)){
          i<-i+1
          next
        }else if(length(d$get(fr))<2){
          i<-i+1
          next
          ### FOR PART 1 ONLY, if bot X has the two chips I'm looking for, return that bot name
        } else if(all(d$get(fr)==c(17,61))){
          return(fr)
          break
        }else{
          ### get Y  - if it's empty, add the low chip.
          ### if it's not empty, add the low chip  to it & sort the order so low chip is first.
          lo<-str_flatten(c(inst[6],inst[7])," ")
          if(d$has(lo)){d$set(lo,sort(c(d$get(lo),d$get(fr)[1])))}else{d$set(lo,d$get(fr)[1])}
          ### get Z - if it's empty, add the high chip.
          ### if not add the high chip and its chips. 
          hi<-str_flatten(c(inst[11],inst[12])," ")
          if(d$has(hi)){d$set(hi,sort(c(d$get(hi),d$get(fr)[2])))}else{d$set(hi,d$get(fr)[2])}
          ### take the chips from bot X
          d$remove(fr)
          ### remove this instruction
          instlist<-instlist[-i]}}}}
  ### when all instructions have been executed, return the dictionary (if the bot hadn't been returned earlier)
  d}
```


```{r}
part1<-microrunner(input)
part1$as_list()[order(unlist(part1$keys()))]
```
## Part 2
Just let it run, until there are no instructions left:

```{r}

microrunner2<-function(instlist){
  d<-dict()
  while(length(instlist)>0){
    i<-1
    while(i<=length(instlist)){
      inst<-instlist[i]
      inst<-unlist(str_split(inst," "))
      if(length(inst)==6){
        adval<-str_flatten(c(inst[5],inst[6])," ")
        if(d$has(adval)){d$set(adval,sort(c(d$get(adval),as.numeric(inst[2]))))}else{d$set(adval,c(as.numeric(inst[2])))}
        instlist<-instlist[-i]
        next
      }else{
        fr<-str_flatten(c(inst[1],inst[2])," ")
        if(!d$has(fr)){
          i<-i+1
          next
        }else if(length(d$get(fr))<2){
          i<-i+1
          next
        } else{
          lo<-str_flatten(c(inst[6],inst[7])," ")
          if(d$has(lo)){d$set(lo,sort(c(d$get(lo),d$get(fr)[1])))}else{d$set(lo,d$get(fr)[1])}
          hi<-str_flatten(c(inst[11],inst[12])," ")
          if(d$has(hi)){d$set(hi,sort(c(d$get(hi),d$get(fr)[2])))}else{d$set(hi,d$get(fr)[2])}
          d$remove(fr)
          instlist<-instlist[-i]}}}}
  d}
```


```{r}
p2<-microrunner2(input)
part2<-c(p2$get("output 0"),p2$get("output 1"),p2$get("output 2"))
part2
part2<-prod(part2)
part2
```

