---
title: "Day 19 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(numbers)
options(scipen = 999)
```

## Part 1
I think I can do this by removing every other elf (if there's an even number of elves) or removing every other elf & then removing first elf, if there's an odd number.

```{r}
### start by creating the elves
part1<-1:5
### while there are still elves 
while(length(part1)>1){
  ### if there are an even number of elves, then the elves in an "odd" position (first, third, fifth,etc.) will get to steal
  ### the elves in an "even" position will be kicked out
  if(length(part1)%%2==0){part1<-part1[c(TRUE,FALSE)]
  ### if there are an odd number of elves, the elves in an "odd" position will steal,
  ### the elves in an even position will be kicked out
  ### and the final "odd" elf will steal the first elf - setting up for a new round.
  }else{part1<-part1[c(TRUE,FALSE)]
  part1<-part1[-1]}}
part1
```

## Part 2
Start by creating the elves
and a pointer to show which elf is the one doing the theiving.
if there are an even number of elves, remove the elf halfway around the circle.
if there an odd number of elves, remove the elf halfway (round down)
increase the pointer
```{r}
part2<-1:20

slowp2<-function(x){
  x<-1:x
  i<-1
  while(length(x)>1){
    ### the one to remove is half of the circle away - mod by the length of the circle if you coil around to the beginning
    rmv<-floor(i+(length(x)/2))%%length(x)
    ### and because 1 indexed, use length of the circle instead of 0
    if(rmv==0){rmv<-length(x)}
    #cat("circle: ",length(x)," elf: ",x[i],"at: ",i," robbed: ",x[rmv]," at: ",rmv,"\n")
    #cat("*******",paste(part2),"\n")
    ### remove that elf
    x<-x[-rmv]
    ## if the elf removed has a lower number than the elf who stole
    ## then everyone will move up a position and you *don't* increase
    ## the index to get to the next elf.
    ## otherwise, increase the index.
    if(rmv>i){i<-i+1}
    if(i>length(x)){i<-1}
    
    #  if(i==1){cat(paste(x),"\n")}
  }
  x}
```
This takes forever.

There's a pattern as to which elves are left at the end of every round, it is dependent on how many elves there are at the beginning- 

```{r}
lastelfstanding<-function(x){
  x<-1:x
  lx<-length(x)
  while(lx>1){
    ### the number of elves is a multiple of 3
    if(lx%%3==0){x<-x[c(FALSE,FALSE,TRUE)]
    }else{
      ### 4 cases
      ### even number of elves and elves MOD 3 is 2
      if(lx%%3==2&&lx%%2==0){
        x<-c(x[1:(lx/2)][c(F,T,F)],x[(lx/2):lx][c(T,F,F)])
        if(any(is.na(x))){x<-x[-(which(is.na(x)))]}
        ###odd number of elves and elves MOD 3 is 2
      }else if(lx%%3==2&&lx%%2==1){
        x<-c(x[1:floor(lx/2)][c(F,T,F)],x[ceiling(lx/2):lx][c(F,T,F)])
        #    cat(x,"\n")
        if(any(is.na(x))){x<-x[-(which(is.na(x)))]}
        ###even number of elves and elves MOD 3 is 1
      }else if(lx%%3==1&&lx%%2==0){
        x<-c(x[1:(lx/2)][c(T,F,F)],x[(lx/2):lx][c(T,F,F)])
        if(any(is.na(x))){x<-x[-(which(is.na(x)))]}
        ###odd number of elves and elves MOD 3 is 1
      }else if(lx%%3==1&&lx%%2==1){
        x<-c(x[1:floor(lx/2)][c(T,F,F)],x[floor(lx/2)],x[ceiling(lx/2):lx][c(F,T,F)])
        if(any(is.na(x))){x<-x[-(which(is.na(x)))]}
      }else{cat("something went wrong")}}
    lx<-length(x)}
  x}
```

```{r}

part2<-lastelfstanding(5)
part2
```
