---
title: "Day XX Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
options(scipen = 999)
```

## Part 1
```{r}
input<-read_lines("Day4Sample.txt")
```

In the most convoluted way possible - 
Split out the string into encryptedname, id, checksum
split the encrypted name into individual pieces
use table to tally (and sort) in descending order
grab the first 5 (which will be alphabetical order, because table does that)
compare to the checksum

```{r}
realroom<-function(rm){
  chksum<-str_sub(rm,-6,-2)
  encrnm<-unlist(str_split(str_sub(rm,end=-8),"-"))
  id<-as.numeric(tail(encrnm,1))
  encrnm<-encrnm[-length(encrnm)]
  encrnm<-unlist(str_split(str_flatten(encrnm),""))
  chkagainst<-table(encrnm)
  chkagainst<-chkagainst[order(chkagainst,decreasing=TRUE)]
  chkagainst<-str_flatten(names(chkagainst[1:5]))
  if(chksum==chkagainst){unname(id)}else{0}}
```


```{r}
part1<-sapply(input,function(x) realroom(x))
part1
```
## Part 2
do the function that allows cipher shift - 

```{r}
shiftit<-function(str,id){
  answer<-""
  for(i in 1:nchar(str)){
    x<-str_sub(str,i,i)
    if(x=="-"){answer<-paste0(answer," ")
    }else{
      x<-(which(x==letters)+id)%%26
      if(x==0){x<-26}
      y<-letters[x]
      answer<-paste0(answer,y)}}
answer}
```

apply it to the input:
```{r}
part2<-as.data.frame(matrix(ncol=1,nrow=0))
for(i in 1:length(input)){
  part2<-rbind(part2,c(input[i]))}
colnames(part2)<-c("room")



part2<-part2 %>% rowwise %>%
  ### get ids
  mutate(ids=realroom(room)) %>%
  ### filter out decoy rooms
  filter(ids!=0) %>%
  ### get the room names
  mutate(roomonly=str_replace(str_sub(room,end=-8),paste0("-",ids),"")) %>%
  ### decrypt them
  mutate(decrypt=shiftit(roomonly,ids))
```
The room names are fun - but so as to not use real input, I'm going to not show them. This line does, though

```{r}
part2$decrypt
```

for real input, search for north.  For this fake, search for "very" just to show how it works
```{r}
part2<-part2$ids[which(str_detect(part2$decrypt,"very"))]
### for real input search for "north"
#part2<-part2$ids[which(str_detect(part2$decrypt,"north"))]
part2
```

